---
title: "FRNL CS 2.10"
output: html_notebook
---

# Evaluation of fodder beet + catch crop rotation (FRNL)

Table 1. Summary of observed and simulated data
```{r, echo=FALSE, include=FALSE}
library(dplyr) # data analysis
library(ggplot2)
library(lubridate) # taking care of dates
library (hydroGOF) # library with heaps of functions that do stats we need
library(xtable)
library(knitr)
library(tidyr)
library(RSQLite)
library(here)
```

```{r, echo=FALSE}
rootPath <- "C:/GitHubRepos/2017_FodderBeetOats/FodderBeatOats"
```

```{r, echo=FALSE, include=FALSE}
# read simulated data
simData <- read.csv(paste0(rootPath,"/SimulatedDataWorked.csv"), header = TRUE)
str(simData)
```


```{r, echo=FALSE, include=FALSE}
# Tidy up sim data
simData_work <- simData %>%
  mutate(Nit=factor(Nit)) %>%
  mutate(Date=as.Date(ymd_hms(Date))) %>%
  mutate(Irr = factor(Irr, levels=c("Nil", "Full"), labels=c("Dry","Irrig")))  %>%
  mutate() %>%
  dplyr::select(-Zone,-SimulationID,-SimName) %>%
  tidyr::gather("VarName","SimValue",Stage_FB:N_leach_prof) %>%
  mutate(VarName = factor(VarName)) %>%
  rowwise() %>%
  mutate(SimValue = ifelse(grepl('_N_Perc_', VarName),SimValue*100,SimValue))

summary(simData_work)
```
```{r, echo=FALSE, include=FALSE}
str(simData_work)
```

```{r, echo=FALSE, include=FALSE}
# read observed data
obsData <- read.csv(paste0(rootPath,"/ObservedDataWorked.csv"), header = TRUE)

```


```{r, echo=FALSE, include=FALSE}
# Average observed data

# Prepare merge variables

# Tidy up

obsData_av <- obsData %>%
  mutate(Nit=factor(Nit)) %>%
  mutate(Date=ymd(Date)) %>%
  dplyr::select(-Plot,-Rep) %>%
  group_by(Date,Irr,Nit) %>%
  summarise_all(funs(mean)) %>%
  tidyr::gather("VarName","ObsValue",Leaf_DM_FB:LAI_FB) %>%
  mutate(VarName = factor(VarName))%>%
  ungroup()

summary(obsData_av)
```

```{r, echo=FALSE, include=FALSE}
str(obsData_av)
```
```{r, echo=FALSE, include=FALSE}
str(simData_work)
```

```{r, echo=FALSE}
mergedDF <- data.frame() # empty data frame to do stuff
mergedDF <- merge(obsData_av, simData_work, by=c("Date","Irr","Nit","VarName"))

mergedDF %>%
gather("VarType","VarValue",ObsValue:SimValue) %>%
mutate(VarType=factor(VarType))%>%
spread(VarName,VarValue) %>%
dplyr::select(-Date) %>%
summary() 
```

# Model evaluation

## 1. Simulated x Observed graphs

- Black line is x=y (1:1 line)
- Colour lines are linear fits for dryland and irrigated crops

```{r, fig.width=12, fig.height=12, echo=FALSE}

mergedDF %>%
  ggplot(aes(x=ObsValue,y=SimValue)) +
  geom_point(alpha=0.5, aes(colour=Irr,shape=Nit), size=2) +
  geom_abline(intercept = 0, slope = 1, size=1.5) +
  facet_wrap(~VarName, ncol=3, scales='free')+
  geom_smooth(method = "lm", se = TRUE,  aes(fill = Irr, linetype=Irr, colour=Irr), alpha=0.1) +
  coord_fixed(ratio = 1) +
  ylab("Simulated value") +
  xlab("Measured value")


```

## 2. Basic statistics

```{r, include=FALSE, echo=FALSE}

# Create stats function
gauchStats <- function(sim, meas) {

  n_s <- length(sim)
  n_m <- length(meas)
  model <- lm(meas~sim)
  sim_sq <- sum((sim - mean(sim))^2)
  mes_sq <- sum((meas - mean(meas))^2)
  r2 <- summary(model)$r.squared
  slope <- model$coefficients[[2]]

  sb <- (sum(mean(meas)) - sum(mean(sim)))^2
  nu <- (1-slope)^2 * (sim_sq/n_s)
  lc <- (1-r2) * (mes_sq/n_m)
  msd <- sb+nu+lc

  sb_r <- round((sb/msd)*100,1)
  nu_r <- round((nu/msd)*100,1)
  lc_r <- round((lc/msd)*100,1)

  msd_r <- sb_r+nu_r+lc_r

  # select which variables to output
  out <- c(sb_r,nu_r,lc_r, msd_r, round(r2*100,1))

  return(out)

}
```



```{r}
# Embed gauch stats in another function 

doStats <-  function (x) {
    library(dplyr)
    x %>% dplyr::summarise(
    n = n(),
    meanObsValue = mean(ObsValue),
    r2 = gauchStats(SimValue,ObsValue)[5],
    rmse = rmse(SimValue,ObsValue),
    r_rmse = rmse(SimValue,ObsValue)/mean(ObsValue)*100, # big precision to enable cretrial of selection
    nse = NSE(SimValue,ObsValue),
    sb = gauchStats(SimValue,ObsValue)[1],
    nu = gauchStats(SimValue,ObsValue)[2],
    lc = gauchStats(SimValue,ObsValue)[3],
    rsr = rsr(SimValue,ObsValue), # Ratio of RMSE to the standard deviation of the ObsValueervations
    pBias = pbias(SimValue,ObsValue)) # Percent Bias between sim and ObsValue, with treatment of missing values.
}
```



Our criteria of a good model performance (mock values as e.g. to discuss):

- R2 > 75%
- RMSE < 20%
- NSE of 1.0
- LC = 100%


```{r, fig.width=12, results='asis',results='asis', echo=FALSE}

mergedDF %>%
  group_by(VarName) %>%
  doStats()

```

## 3. Sensibility analysis

- Criteria: Are simulated values sensible in relation to real life experiences?

```{r, fig.height = 8, echo=FALSE}
simData_work %>%
  ungroup()%>%
  filter(VarName == "Drainage"|VarName == "N_leach_prof"|VarName == "N_min_prof") %>%
  group_by(VarName, Irr, Nit) %>%
  arrange(VarName, Irr, Nit,Date) %>% 
  mutate(SimValue=cumsum(SimValue)) %>%
  ggplot(aes(x=Date,y=SimValue)) +
  geom_line(alpha=0.5, aes(colour=Irr,linetype=Nit)) +
  facet_grid(VarName~.,  scales='free')
```


## 4. Sensitivity analysis

- Placeholder for possible sensitivity analysis in APSIM-X
- Which parameters to test?

```{r}

```

