---
title: "R Notebook"
output:
  word_document: default
  html_notebook: default
---

Inspired by Hamish's beautiful Python script

AIM: 
- Read pre-processed observed data
- Save as APSIM-X format
```{r}
library(readxl)
library(lubridate)# taking care of dates
library(dplyr) # data analysis
library(ggplot2)
library(xlsx)

```

## File locations

```{r}
# where child observed data in APSIM-X format is READ from common K: drive
pathObs <- "K:/Rainshelter/Fodderbeet 2016-2017/Analysis/obsFiles_ForAPSIM" 

# where master observed data will be SAVED TO
pathMaster <- "K:/Rainshelter/Fodderbeet 2016-2017/Analysis" 

```

## Define name convention as per APSIM simulations

- lookup take 

```{r}
labelNames <- read.table(paste0(pathMaster,"/labels.txt"), header=TRUE)
labelNames
```



## Which files where found?
```{r}
obs.files <- list.files(pathObs,pattern="\\.xlsx$", full.names=FALSE) # if true gets path
print(obs.files)
```

##Read and compile files

- Creates a single observed data file

```{r}

for (i in 1:length(obs.files)) {

# FIXME: add a TRYCATCH here that continues the loop in the case file does not comply with standard, error message with problem files
# FIXME: sheet name or columns name
# FIXME: Set up name convention for files
  
this.obs.file <- xlsx::read.xlsx(paste0(pathObs,"\\",obs.files[i]), sheetName = "OBS")  

this.obs.file$SimulationName <- factor(this.obs.file$SimulationName)
this.obs.file$Clock.Today <- ymd(this.obs.file$Clock.Today)


#ymd_hms

#print(head(this.obs.file))

  if(i==1) {
  
  obs.master <- data.frame()
  obs.master <- this.obs.file
  
  } else {
  
  # Two compulsory fields are "SimulationName" and "Clock.Today"
  obs.master <- merge(obs.master,this.obs.file, by=c("SimulationName","Clock.Today"), all=TRUE)
 
  }

}

summary(obs.master)
```

```{r}
str(obs.master)
summary(obs.master)
```

## Save

- As single  observed file

```{r}

df1 <- merge(obs.master,labelNames, by="SimulationName")

# tidy up data
df2 <- df1 %>%
  mutate(Clock.Today=as.character(as.Date(Clock.Today, "%Y-%m-%d"))) %>% # FIXME: problems in data format
  arrange(Clock.Today) %>%
  as.data.frame() 

summary(df2)
```
For the git repo in local
```{r}
# save in common K drive folder
write.csv(df2,paste0(pathMaster,"/OBS.csv"), na ="", row.names = FALSE, quote = FALSE)
print(paste0("Save file to: ", pathMaster))
```

# Graph for FRNL presentation

FIXME: Move this elsewhere
```{r}
df2 %>%
  mutate(Irr=factor(Irr, levels=c("Nil","Full"),labels=c("Dryland","Irrigated"))) %>%
  dplyr::select(Clock.Today,Nit, Irr,
                FodderBeet.AboveGround.Live.Wt) %>%
  ggplot(aes(x=ymd(Clock.Today),
             y=FodderBeet.AboveGround.Live.Wt*10/1000, colour=factor(Irr),linetype=factor(Nit)
             )) +
  geom_point(size=3, alpha=0.5,aes(shape=factor(Nit))) +
  geom_smooth() +
  ylab("Above ground biomass \n(t DM/ha)") +
  xlab("Date") +
  labs(shape="Nitrogen \nfertiliser (kg/ha)", colour="Irrigation") + 
  guides(linetype=FALSE) + # remove legend
  theme(text = element_text(size=16))
#+ theme(legend.position="top")
```

