---
title: "R Notebook"
output: html_notebook
---
##AIM:
Reads raw or pre-processes raw observed data and creates two outputs:

(i) Observed data for report and analysis (i.e. all variables listed in a "Variables" column)
(ii) An observed file for APSIM-X format (i.e. each variable is a column)

Change units to match APSIM-X units of each variable

```{r}
library(lubridate)
library(ggplot2)
library(dplyr)
library(tidyr)
library(readxl)
library(xlsx)
```

```{r}
# dir address
pathFolderRaw <- "K:/Rainshelter/Fodderbeet 2016-2017/Analysis/obsFiles_PreProc" # where to read from
print(paste0("Read data from: ", pathFolderRaw))

nameAddressOfFile <- paste0(pathFolderRaw, "/DataFrame_FodderBeetBiomass.xlsx")
```

```{r}
pathFolderWork <- "K:/Rainshelter/Fodderbeet 2016-2017/Analysis/obsFiles_ForAPSIM"  # where to save to
print(paste0("Save data to: ", pathFolderWork))
```


```{r}
obsDataRaw <- read_xlsx(nameAddressOfFile, sheet = "DataFrame", col_names = TRUE) # I find this synthax more transparent - without the risk of reading wrong sheet - eit

#obsDataRaw <- xlsx::read.xlsx(nameAddressOfFile, sheetIndex = 4, header = TRUE, startRow = 4) # FIXME: why this synthax with less clear parameters?

# remove NA rows 
ind <- apply(obsDataRaw, 1, function(obsDataRaw) all(is.na(obsDataRaw)))
obsDataRaw <- obsDataRaw[ !ind, ]

 
summary(obsDataRaw)
```

# Get units
```{r}
dfTemp <- read_xlsx(nameAddressOfFile, sheet = "lookupVariables", col_names = TRUE)
# obsDataRaw <- xlsx::read.xlsx(nameAddressOfFile, sheetIndex = 4, header = TRUE, startRow = 4 )

# remove NA rows 
ind <- apply(dfTemp, 1, function(dfTemp) all(is.na(dfTemp)))
obsDataUnits <- dfTemp[ !ind, ]

# for all cols to factors
cols <- colnames(obsDataUnits)
obsDataUnits[cols] <- lapply(obsDataUnits[cols], factor)

dfTemp <- NULL
 
head(obsDataUnits, nrow(obsDataUnits))
```
# Create consistent data

- FIXME: Mei to tidy up
```{r}
obsDataConsist <- obsDataRaw %>%
#  dplyr::select(-SimulationName, -Code) %>%
  gather("VarNameAPSIM","VarValue",FodderBeet.Leaf.Live.Wt:FodderBeet.Leaf.LAI) %>%
  mutate(VarNameAPSIM=factor(VarNameAPSIM)) %>%
  merge(obsDataUnits, by.x="VarNameAPSIM",by.y="Name_APSIM")

summary(obsDataConsist)




```
## Save consistent data
```{r}
# save consistent data
write.csv(obsDataConsist, paste0("K:/Rainshelter/Fodderbeet 2016-2017/Analysis/consistentData/CONS_Biomass_FodderBeet.csv"), row.names = FALSE)
```

## Prepare data for APSIM

- Select variables
- Correct units 
```{r}

# Biomass units from kg/ha to g/m2 and N conc units from % to 0-1
  obsData_APSIM <-  obsDataConsist %>%
  dplyr::select(Clock.Today,SimulationName, VarNameAPSIM,VarValue) %>%
  mutate(VarValue = ifelse(grepl('.Wt', VarNameAPSIM),VarValue/10,VarValue)) %>% 
  mutate(VarValue = ifelse(grepl('.NConc', VarNameAPSIM), VarValue/100,VarValue)) %>% 
  mutate(Clock.Today = ymd(Clock.Today)) %>%
  mutate(SimulationName=factor(SimulationName)) %>%
  # do averages for APSIM
  group_by(Clock.Today,SimulationName,VarNameAPSIM) %>%
  summarise(VarValue=as.numeric(mean(VarValue))) %>%
  # spread vars as per apasim format
  spread(VarNameAPSIM,VarValue) %>%
  as.data.frame() # Format needed to save as excel

head(obsData_APSIM)
```
## Add new variables 

- Create variables as per APSIM-X test neeeded
- FIXME: This chunck is UNTIDY as is - let's fix it 
- FIXME: Could be other script reading consistent data (?)
```{r}
 # Total N by organ (FIXME: find simpler way of doing that)
 obsData_APSIM$FodderBeet.Leaf.N <- obsData_APSIM$FodderBeet.Leaf.Live.Wt*obsData_APSIM$FodderBeet.Leaf.NConc

 obsData_APSIM$FodderBeet.Petiole.N <- obsData_APSIM$FodderBeet.Petiole.Live.Wt*obsData_APSIM$FodderBeet.Petiole.NConc
 
 obsData_APSIM$FodderBeet.StorageRoot.N <- obsData_APSIM$FodderBeet.StorageRoot.Live.Wt*obsData_APSIM$FodderBeet.StorageRoot.NConc
 
 # Total biomass FIXME: not nice to overwhite here ... find better way to do it
  obsData_APSIM$FodderBeet.AboveGround.Live.Wt <- obsData_APSIM %>% dplyr::select(ends_with(".Live.Wt")) %>% rowSums() # FIXME: Some bug here
 
 
# Sort out factors
# factorindex <- c("Nit","Irr","Plot","Rep")
# obsData_APSIM[,factorindex] <- as.data.frame(sapply(obsData_APSIM[,factorindex],as.factor))
#work well if using index <- c(2:3,4,5...)
```

## Check the data

```{r}
str(obsData_APSIM)
```


```{r}
head(obsData_APSIM)
```

```{r}
summary(obsData_APSIM)
```

## save for APSIM-X
```{r}
write.xlsx(obsData_APSIM,paste0(pathFolderWork,"/OBS_Biomass_FodderBeet.xlsx"), sheetName="OBS", row.names = FALSE)
```

