---
title: "R Notebook"
output: html_notebook
---
##AIM:
Reads raw or pre-processes raw observed data and creates two outputs:

(i) Observed data for report and analysis (i.e. all variables listed in a "Variables" column)
(ii) An observed file for APSIM-X format (i.e. each variable is a column)

Change units to match APSIM-X units of each variable

```{r, warning = FALSE}
library(lubridate)
library(ggplot2)
library(dplyr)
library(tidyr)
library(readxl)
library(xlsx)
```

```{r}

# dir address

pathFolderRaw <- "K:/Rainshelter/Fodderbeet 2016-2017/Analysis/obsFiles_PreProc" # where to read from
print(paste0("Read data from: ", pathFolderRaw))

nameAddressOfFile <- paste0(pathFolderRaw, "/DataFrame_FodderBeetBiomass.xlsx")



```

```{r}
pathFolderWork <- "K:/Rainshelter/Fodderbeet 2016-2017/Analysis/obsFiles_ForAPSIM"  # where to save to
print(paste0("Save data to: ", pathFolderWork))
```
```{r}
obsDataRaw <- read_xlsx(nameAddressOfFile, sheet = "DataFrame", col_names = TRUE)
 
summary(obsDataRaw)
```


## Tidy up data
```{r}

# Biomass units from kg/ha to g/m2
df_temp <-  obsDataRaw %>%
  tidyr::gather("VarName","VarValue", FodderBeet.Leaf.Live.Wt:FodderBeet.Leaf.LAI) %>% 
  mutate(VarName=factor(VarName)) %>%
  mutate(VarValue = ifelse(grepl('.Wt', VarName),VarValue/10,VarValue)) %>% # 
  tidyr::spread(VarName,VarValue)

 # FIXME: get name right here
 df_temp$FodderBeet.AboveGround.Live.Wt <- df_temp %>% select(ends_with(".Live.Wt")) %>% rowSums()

# dates
obsData <- df_temp %>% mutate(Clock.Today = ymd(Clock.Today)) 
 

# Sort out factors
factorindex <- c("Nit","Irr","Plot","Rep", "Code")
obsData[,factorindex] <- as.data.frame(sapply(obsData[,factorindex],as.factor))
#work well if using index <- c(2:3,4,5...)
```

## Check the data

```{r}
str(obsData)
```


```{r}
head(obsData)
```

```{r}
summary(obsData)
```
Graph

```{r, fig.width = 12}
obsData %>%
  ggplot(aes(x=Clock.Today,y=(FodderBeet.Leaf.Live.Wt),colour=Irr, linetype=Nit,shape=Nit)) +
  geom_point(size=2,alpha=0.5) +
  geom_line() +
  ylab("g/m2") +
  facet_grid(.~Rep)
```

```{r}
obsData %>%
  ggplot(aes(x=Clock.Today,y=(FodderBeet.Leaf.NConc),colour=Irr, linetype=Nit,shape=Nit)) +
  geom_point(size=2,alpha=0.5) +
  geom_line() +
  ylab("%") +
  facet_grid(.~Rep)
```
## Compare N% and biomass in a organ
```{r}
obsData %>%
  ggplot(aes(x=FodderBeet.Leaf.Live.Wt,y=FodderBeet.Leaf.NConc,colour=Irr, shape=Nit)) +
  geom_point(size=2,alpha=0.5) +
#  geom_smooth(aes(linetype=Nit)) +
  ylab("N%") +
  xlab("g DM/m2") +
  facet_grid(.~Rep)
```

## Save observed data

```{r}
head(obsData)

```

- For APSIM-X
```{r}
# FIXME: Rob to add variables and fix units as needed 
df <- obsData %>%
 # mutate(Irr=factor(Irr),Nit=factor(Nit)) %>%
  mutate(Clock.Today=as.Date(ymd(Clock.Today))) %>%
  mutate(Clock.Today=as.character(as.Date(Clock.Today, "%Y-%m-%d"))) %>%
  dplyr::select(-Plot,-Rep,-Nit,-Irr,-Code) %>%
  group_by(Clock.Today,SimulationName) %>%
  summarise_all(funs(mean)) %>%
  mutate(SimulationName = factor(SimulationName))%>%
  ungroup()%>%
  as.data.frame()

summary(df)

write.xlsx(df,paste0(pathFolderWork,"/OBS_Biomass.xlsx"), sheetName="OBS", row.names = FALSE) 
```


